---
title: "petite_clean_data"
output: html_document
date: '2023-04-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(cowplot)
#library(ggblur)
# set wd
library(multcomp)
library(agricolae)
setwd("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data/depth/10_DepthAnalyses/")
source("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data/bioscreen_functions.R")
library(viridis)
setwd("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data/raw_outcross_petite/")

# file with metadata of bioscreen wells
codes <- read.csv("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data/raw_outcross_petite/bioscreen20230418BIO1_codes.csv") # 
biosc2 <- read.csv("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data/raw_outcross_petite/20230418_BIO2.csv")

```

Setting color palette
```{r}
relfitPalette <- c("#787878", "#FF9A17", "#3B83FF","#3EFF3B", "#FD73FF", "#FFF200", "#A880FF")
relfitPalette2 <- c("#FF9A17", "#3B83FF","#3EFF3B", "#FD73FF", "#FFF200", "#A880FF")
petitePalette2 = c("white", "black")

legend_title1 <- "Test\nEnvironment"
legend_title2 <- "Petite"
```

Combine dataframes for lines from outcrossing experiment

```{r}

# Transpose dataframe
Time <- biosc2$Time #remember row id
bio2 <- as.data.frame(t(biosc2[,-1])) #transpose dataframe, exclude 1st colunmn
colnames(bio2) <- Time #create column names  

# Combine the bioscreen output with metadata
df = cbind(codes, bio2)


```

Plot petite outcross growth graphs

```{r}

# loop through matrix to get growth curves, maximum slope and final OD per well
pdf("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data/figures/outcross_petite_metals.pdf", onefile = TRUE)
par(mfrow = c(10,4))
par(mar = c(0,0,0,0))
par(cex=0.5, mai=c(0.1,0.2,0.1,0.1))
for(r in 1:nrow(df)){
  temp = as.numeric(df[r, 7:197])
  time = seq(0.25, length(temp)/4, by = 0.25)
  growth <- spline.slope(time, temp)
  t <- spline(time, temp)
   codes$initOD[r] <- mean(temp[3:8]) # Arbitrary choice, first between 45 min and 2.5 hours
   # codes$OD_maxSlope[r] <- temp[which(time == t)]
    codes$finalOD[r] <- mean(temp[(length(temp)-5):length(temp)])
    codes$maxSlope[r] <- growth
    codes$time_maxSlope[r] <- t
    codes$finalTime[r] <- length(temp)/4
    plot(temp ~ time, xlim = c(0, 70), ylim = c(0, 1.5), xlab = "Time (h)", ylab = "OD")
    text(5, 1, paste0(codes$metal[r], " ", codes$strain[r], "\nC ", codes$conc[r],"\nr = ", round(growth, 3), "\ntime points\nlost: "), cex = 0.5)
     abline(v=t)
  }

dev.off()


```

Get relative fitness data to merge with the coverage files
```{r}
out_mydata <- codes %>% 
  dplyr::rename(
     E2 = conc
    )
```

Clean dataframe
```{r}
out_OLY077_data <- out_mydata %>% 
  filter(strain=="OLY077" ) %>% 
  arrange(E2) %>% 
  select(strain, E2, maxSlope)
  

```
out_OLY077_data <- out_OLY077_data %>% 
  dplyr::group_by(E2) %>% 
  summarise(anc_slope = mean(maxSlope))

```{r}
out_mydata3 <- merge(out_OLY077_data, out_mydata, by = 'E2')

out_mydataOLY <- out_mydata3 %>% 
  filter(strain.y=="OLY077")

out_mydata3 <- out_mydata3 %>% 
  mutate(relfit = maxSlope.y - maxSlope.x)

out_mydata3 <- out_mydata3 %>% 
  mutate(E2 = case_when(E2 == "YPAD" ~ ".YPAD",
                        E2 == "CO" ~ "Co",
                        E2 == "MN" ~ "Mn",
                        )) 
```

Load metal experimental data 

```{r}
setwd("~/Documents/GitHub/crossTolerance_bazzicalupoEtAl/cross_tol_data")
name_codes = read_csv("code_strain_line.csv")

mydata6 = read.csv("mbm_intermediate_files/mydata6.csv")

out_mydata6 <- out_mydata3 %>% 
  select(strain.y, E2,  maxSlope.y, maxSlope.x) %>% 
  arrange(strain.y, E2) %>% 
  group_by(strain.y, E2, maxSlope.x) %>% 
  dplyr::summarise(avg_slope = mean(maxSlope.y)) %>% 
  mutate(relfit1 = avg_slope - maxSlope.x)

out_mydata6[out_mydata6 == "OLY077"] <- "W303"
mydata6[mydata6 == "blank"] <- "Blank"

out_mydata6 <- out_mydata6 %>% 
  dplyr::filter(strain.y != "Blank")

```

Add petite phenotype information

```{r}
table(out_petite$petite)
table(petites$petite)

out_petites = read_csv("raw_outcross_petite/outcross_petite.csv")
legend_title_none = ""
out_mydata6$strain <- out_mydata6$strain.y
out_petite <- merge(out_mydata6,out_petites,by="strain")


out_summary1 <- out_petite %>% 
  #filter(E2!=".YPAD") %>% 
  Rmisc::summarySE(measurevar = "relfit1", groupvars = c( "E2", "petite"))

out_pet=out_petite %>%
  na.omit() %>% 
  #filter(E2!=".YPAD") %>% 
  ggplot(aes(x=E2,y=relfit1,fill=petite)) +
    geom_point(alpha=0.2, size=0.5, position=position_dodge(width=0.75),aes(group=petite)) + 
    stat_summary(fun=mean, geom="point", shape=21, size=4,
             position=position_dodge2(width=0.75, preserve="single")) +
    geom_errorbar(data=out_summary1, aes(ymin = relfit1 - se, ymax = relfit1 + se), width=0.75, position=position_dodge2(width=0.75, preserve="single")) +
     # scale_fill_discrete(limits = c("np", "p"),labels = c("Grande", "Petite"), values=petitePalette2)+
    scale_fill_manual(legend_title_none, limits = c("np", "p"),labels = c("Grande", "Petite"), values=petitePalette2) +
    labs(x="Test Environment", y="") +
    theme_bw() +   geom_hline(yintercept=0, size=1)+
    theme(legend.title=element_text(size=rel(0.75))) +
    geom_vline(xintercept = 1.5) + 
    geom_vline(xintercept = 2.5) + 
    geom_vline(xintercept = 3.5) + 
    geom_vline(xintercept = 4.5) + 
    geom_vline(xintercept = 5.5) +
    geom_vline(xintercept = 6.5)
#ggsave("figures/petite_outcross.jpeg", width=8, height=6, units = "in", dpi=1000)




```
oucrossing study Petite t-test analysis out_petite 
```{r}
all_res_aov <- aov(relfit1 ~ petite*E2,
  data = out_petite
)

summary(all_res_aov)




```
Petite analysis
```{r}
name_codes = read_csv("code_strain_line.csv")
petites = read_csv("raw_outcross_petite/petites.csv")
colnames(petites)[colnames(petites) == "evoline"] ="code_0"
petites1 <- merge(petites,name_codes,by="code_0")
colnames(petites1)[colnames(petites1) == "code_2"] ="strain"
mydata7 <- merge(mydata6,petites1,by="strain")
pet_summary1 <- mydata7 %>% 
  Rmisc::summarySE(measurevar = "relfit1", groupvars = c( "E2", "petite"))

met_pet=mydata7 %>%
  na.omit() %>% 
  ggplot(aes(x=E2,y=relfit1,fill=petite)) +
    geom_point(alpha=0.2, size=0.5, position=position_dodge(width=0.75),aes(group=petite)) + 
    stat_summary(fun=mean, geom="point", shape=21, size=4,
             position=position_dodge2(width=0.75, preserve="single")) +
    geom_errorbar(data=pet_summary1, aes(ymin = relfit1 - se, ymax = relfit1 + se), width=0.75, position=position_dodge2(width=0.75, preserve="single")) +
    scale_fill_manual(values=petitePalette2) +
    #scale_fill_manual(legend_title_none, limits = c("Non-petite (n = 70)", "Petite (n =38)"),labels = c("Grande (n = 70)", "Petite (n = 38)"), values=petitePalette2) +
    labs(x="Test Environment", y="Relative Fitness") +
    theme_bw() +  geom_hline(yintercept=0, size=1)+
    theme(legend.title=element_text(size=rel(0.75))) +
    geom_vline(xintercept = 1.5) + 
    geom_vline(xintercept = 2.5) + 
    geom_vline(xintercept = 3.5) + 
    geom_vline(xintercept = 4.5) + 
    geom_vline(xintercept = 5.5) +
    geom_vline(xintercept = 6.5)
#ggsave("figures/petite_test_environment.jpeg", width=8, height=6, units = "in", dpi=1000)
```

This study Petite t-test analysis
```{r}
our_res_aov <- aov(relfit1 ~ petite*E2,
  data = mydata7
)

summary(our_res_aov)

post_test <- TukeyHSD(our_res_aov)

post_test

summary(post_test)

t.test(relfit1 ~ petite, data = mydata7)

petite_cd = mydata7 %>%
  filter(E2==c("Cd",".YPAD")) 
t.test(relfit1 ~ petite, data = petite_cd)
cd_res_aov <- aov(relfit1 ~ petite*E2,
  data = petite_cd
)
summary(cd_res_aov)

petite_co = mydata7 %>%
  filter(E2==c("Co",".YPAD")) 
t.test(relfit1 ~ petite, data = petite_co)
co_res_aov <- aov(relfit1 ~ petite*E2,
  data = petite_co
)
summary(co_res_aov)

petite_cu = mydata7 %>%
  filter(E2==c("Cu",".YPAD")) 
t.test(relfit1 ~ petite, data = petite_cu)
cu_res_aov <- aov(relfit1 ~ petite*E2,
  data = petite_cu
)
summary(cu_res_aov)


petite_mn = mydata7 %>%
  filter(E2==c("Mn",".YPAD")) 
t.test(relfit1 ~ petite, data = petite_mn)
mn_res_aov <- aov(relfit1 ~ petite*E2,
  data = petite_mn
)
summary(mn_res_aov)


petite_ni = mydata7 %>%
  filter(E2==c("Ni",".YPAD")) 
t.test(relfit1 ~ petite, data = petite_ni)
ni_res_aov <- aov(relfit1 ~ petite*E2,
  data = petite_ni
)
summary(ni_res_aov)


petite_zn = mydata7 %>%
  filter(E2==c("Zn",".YPAD")) 
t.test(relfit1 ~ petite, data = petite_zn)
zn_res_aov <- aov(relfit1 ~ petite*E2,
  data = petite_zn
)
summary(zn_res_aov)


```

Plot all petite performace together
```{r}
# extract legends from the plots
legendout <- get_legend(out_pet)

both_pet=plot_grid(
  met_pet+ theme(legend.position = "none"), out_pet, labels = c("A","B"), rel_widths = c(2, 1.75)
)
ggsave("figures/met_out_petites.jpeg", width=10, height=5, units = "in", dpi=700)
```


Three separate by metal stacked
```{r }
mito_weird = read_csv("raw_outcross_petite/mito_weird100.csv") # windows of 100 bp
petiteCopper = c( "#3EFF3B",  "#787878")
petiteNickel = c("#FFF200","goldenrod1","#787878")
petiteManganese = c("lavenderblush","thistle1","lightpink","lightpink2","palevioletred1","hotpink","deeppink","violetred","deeppink3","hotpink3","lightcoral","violet","orchid3","orchid4","#787878")
petiteAll =c( "#3EFF3B","lavenderblush","thistle1","lightpink","lightpink2","palevioletred1","hotpink","deeppink","violetred","deeppink3","hotpink3","lightcoral","violet","orchid3","orchid4","#FFF200","goldenrod1","#787878")
legend_strain = "Yeast Line"
legend_blank =""

mito_All=mito_weird %>%
  ggplot(aes(x=pos,y=cov)) +
  geom_point(alpha=0.8, size=0.75,aes(colour=strain)) +
  scale_color_manual(legend_blank, values=petiteAll) +
  theme_bw()+
  ylim(0,9000)+
  theme(legend.key.height = unit(0.25, "cm"))+
  annotate("rect", xmin = 6546, xmax = 8194, ymin = 7500, ymax = 8000,
           alpha = .25,fill = "blue", )+annotate("text", x = (6546+8194)/2, y = 8300, label = "15S rRNA", size = 3) +
   annotate("rect", xmin = 13818, xmax = 26701, ymin = 7500, ymax = 8000,#13818	26701	COX1 
           alpha = .25,fill = "blue", )+annotate("text", x = (13818+26701)/2, y = 8300, label = "COX1", size = 3) +
  annotate("rect", xmin = 27666, xmax = 27812, ymin = 7500, ymax = 8000,#27666	27812	ATP8 
           alpha = .25,fill = "blue", )+annotate("text", x = (27666+27812)/2, y = 8800, label = "ATP8", size = 3) +
  annotate("rect", xmin = 28487, xmax = 29266, ymin = 7500, ymax = 8000,#28487	29266	ATP6
           alpha = .25,fill = "blue")+annotate("text", x = (28487+29266)/2, y = 8200, label = "ATP6", size = 3)+
  annotate("rect", xmin = 36540, xmax = 43647, ymin = 7500, ymax = 8000, #36540	43647	COB
           alpha = .25,fill = "blue")+annotate("text", x = (36540+43647)/2, y = 8300, label = "COB", size = 3)+
  annotate("rect", xmin = 46723, xmax = 46953, ymin = 7500, ymax = 8000, #46723	46953	OLI1
           alpha = .25,fill = "blue", )+annotate("text", x = (46723+46953)/2, y = 8800, label = "OLI1", size = 3) +
  annotate("rect", xmin = 48901, xmax = 50097, ymin = 7500, ymax = 8000, #	VAR1
           alpha = .25,fill = "blue")+annotate("text", x = (48901+50097)/2, y = 8200, label = "VAR1", size = 3)+
  annotate("rect", xmin = 58009, xmax = 62447, ymin = 7500, ymax = 8000, #58009	62447	21S_RRNA 
           alpha = .25,fill = "blue", )+annotate("text", x = (58009+62447)/2, y = 8300, label = "21S rRNA", size = 3) +
  annotate("rect", xmin = 73758, xmax = 74513, ymin = 7500, ymax = 8000, #73758	74513	COX2 
           alpha = .25,fill = "blue", )+annotate("text", x = (74513+73758)/2, y = 8300, label = "COX2", size = 3) +
  annotate("rect", xmin = 79213, xmax = 80022, ymin = 7500, ymax = 8000,#79213	80022	COX3 
           alpha = .25,fill = "blue")+annotate("text", x = (79213+80022)/2, y = 8700, label = "COX3", size = 3)+
  annotate("rect", xmin = 85295, xmax = 85777, ymin = 7500, ymax = 8000,#85295	85777	RPM1
           alpha = .25,fill = "blue", )+annotate("text", x = (85295+85777)/2, y = 8300, label = "RPM1", size = 3)+
  labs(x="", y="Mitochondrial coverage", size = 15) 

mito_Mn=mito_weird %>%
  filter(E1 %in% c("Mn", "YPAD")) %>%
  ggplot(aes(x=pos,y=cov)) +
  geom_point(alpha=0.8, size=0.75,aes(colour=strain)) +
  scale_color_manual(legend_blank, values=petiteManganese) +
  theme_bw()+
  ylim(0,9000)+
  theme(legend.key.height = unit(0.25, "cm"))+
  annotate("rect", xmin = 6546, xmax = 8194, ymin = 7500, ymax = 8000,
           alpha = .25,fill = "blue", )+annotate("text", x = (6546+8194)/2, y = 8300, label = "15S rRNA", size = 3) +
   annotate("rect", xmin = 13818, xmax = 26701, ymin = 7500, ymax = 8000,#13818	26701	COX1 
           alpha = .25,fill = "blue", )+annotate("text", x = (13818+26701)/2, y = 8300, label = "COX1", size = 3) +
  annotate("rect", xmin = 27666, xmax = 27812, ymin = 7500, ymax = 8000,#27666	27812	ATP8 
           alpha = .25,fill = "blue", )+annotate("text", x = (27666+27812)/2, y = 8800, label = "ATP8", size = 3) +
  annotate("rect", xmin = 28487, xmax = 29266, ymin = 7500, ymax = 8000,#28487	29266	ATP6
           alpha = .25,fill = "blue")+annotate("text", x = (28487+29266)/2, y = 8200, label = "ATP6", size = 3)+
  annotate("rect", xmin = 36540, xmax = 43647, ymin = 7500, ymax = 8000, #36540	43647	COB
           alpha = .25,fill = "blue")+annotate("text", x = (36540+43647)/2, y = 8300, label = "COB", size = 3)+
  annotate("rect", xmin = 46723, xmax = 46953, ymin = 7500, ymax = 8000, #46723	46953	OLI1
           alpha = .25,fill = "blue", )+annotate("text", x = (46723+46953)/2, y = 8800, label = "OLI1", size = 3) +
  annotate("rect", xmin = 48901, xmax = 50097, ymin = 7500, ymax = 8000, #	VAR1
           alpha = .25,fill = "blue")+annotate("text", x = (48901+50097)/2, y = 8200, label = "VAR1", size = 3)+
  annotate("rect", xmin = 58009, xmax = 62447, ymin = 7500, ymax = 8000, #58009	62447	21S_RRNA 
           alpha = .25,fill = "blue", )+annotate("text", x = (58009+62447)/2, y = 8300, label = "21S rRNA", size = 3) +
  annotate("rect", xmin = 73758, xmax = 74513, ymin = 7500, ymax = 8000, #73758	74513	COX2 
           alpha = .25,fill = "blue", )+annotate("text", x = (74513+73758)/2, y = 8300, label = "COX2", size = 3) +
  annotate("rect", xmin = 79213, xmax = 80022, ymin = 7500, ymax = 8000,#79213	80022	COX3 
           alpha = .25,fill = "blue")+annotate("text", x = (79213+80022)/2, y = 8700, label = "COX3", size = 3)+
  annotate("rect", xmin = 85295, xmax = 85777, ymin = 7500, ymax = 8000,#85295	85777	RPM1
           alpha = .25,fill = "blue", )+annotate("text", x = (85295+85777)/2, y = 8300, label = "RPM1", size = 3)+
  labs(x="", y="Mitochondrial coverage", size = 15)  

#ggsave("figures/mito_weird_cov.jpeg", width=8, height=4, units = "in", dpi=1000)

mito_Ni=mito_weird %>%
  filter(E1 %in% c("Ni", "YPAD")) %>%
  ggplot(aes(x=pos,y=cov)) +
  geom_point(alpha=0.8, size=0.75,aes(colour=strain)) +
  scale_color_manual(legend_blank, values=petiteNickel) +
  theme_bw()+
  ylim(0,9000)+
  theme(legend.key.height = unit(0.25, "cm"))+
  annotate("rect", xmin = 6546, xmax = 8194, ymin = 7500, ymax = 8000,
           alpha = .25,fill = "blue", )+annotate("text", x = (6546+8194)/2, y = 8300, label = "15S rRNA", size = 3) +
   annotate("rect", xmin = 13818, xmax = 26701, ymin = 7500, ymax = 8000,#13818	26701	COX1 
           alpha = .25,fill = "blue", )+annotate("text", x = (13818+26701)/2, y = 8300, label = "COX1", size = 3) +
  annotate("rect", xmin = 27666, xmax = 27812, ymin = 7500, ymax = 8000,#27666	27812	ATP8 
           alpha = .25,fill = "blue", )+annotate("text", x = (27666+27812)/2, y = 8800, label = "ATP8", size = 3) +
  annotate("rect", xmin = 28487, xmax = 29266, ymin = 7500, ymax = 8000,#28487	29266	ATP6
           alpha = .25,fill = "blue")+annotate("text", x = (28487+29266)/2, y = 8200, label = "ATP6", size = 3)+
  annotate("rect", xmin = 36540, xmax = 43647, ymin = 7500, ymax = 8000, #36540	43647	COB
           alpha = .25,fill = "blue")+annotate("text", x = (36540+43647)/2, y = 8300, label = "COB", size = 3)+
  annotate("rect", xmin = 46723, xmax = 46953, ymin = 7500, ymax = 8000, #46723	46953	OLI1
           alpha = .25,fill = "blue", )+annotate("text", x = (46723+46953)/2, y = 8800, label = "OLI1", size = 3) +
  annotate("rect", xmin = 48901, xmax = 50097, ymin = 7500, ymax = 8000, #	VAR1
           alpha = .25,fill = "blue")+annotate("text", x = (48901+50097)/2, y = 8200, label = "VAR1", size = 3)+
  annotate("rect", xmin = 58009, xmax = 62447, ymin = 7500, ymax = 8000, #58009	62447	21S_RRNA 
           alpha = .25,fill = "blue", )+annotate("text", x = (58009+62447)/2, y = 8300, label = "21S rRNA", size = 3) +
  annotate("rect", xmin = 73758, xmax = 74513, ymin = 7500, ymax = 8000, #73758	74513	COX2 
           alpha = .25,fill = "blue", )+annotate("text", x = (74513+73758)/2, y = 8300, label = "COX2", size = 3) +
  annotate("rect", xmin = 79213, xmax = 80022, ymin = 7500, ymax = 8000,#79213	80022	COX3 
           alpha = .25,fill = "blue")+annotate("text", x = (79213+80022)/2, y = 8700, label = "COX3", size = 3)+
  annotate("rect", xmin = 85295, xmax = 85777, ymin = 7500, ymax = 8000,#85295	85777	RPM1
           alpha = .25,fill = "blue", )+annotate("text", x = (85295+85777)/2, y = 8300, label = "RPM1", size = 3)+
  labs(x="Yeast Mitochondrial Reference Genome (bases)", y="", size = 15)  

mito_Cu=mito_weird %>%
  filter(E1 %in% c("Cu", "YPAD")) %>%
  ggplot(aes(x=pos,y=cov)) +
  geom_point(alpha=0.8, size=0.75,aes(colour=strain)) +
  scale_color_manual(legend_strain, values=petiteCopper) +
  theme_bw()+
  ylim(0,9000)+
  theme(legend.key.height = unit(0.25, "cm"))+
  annotate("rect", xmin = 6546, xmax = 8194, ymin = 7500, ymax = 8000,
           alpha = .25,fill = "blue", )+annotate("text", x = (6546+8194)/2, y = 8300, label = "15S rRNA", size = 3) +
   annotate("rect", xmin = 13818, xmax = 26701, ymin = 7500, ymax = 8000,#13818	26701	COX1 
           alpha = .25,fill = "blue", )+annotate("text", x = (13818+26701)/2, y = 8300, label = "COX1", size = 3) +
  annotate("rect", xmin = 27666, xmax = 27812, ymin = 7500, ymax = 8000,#27666	27812	ATP8 
           alpha = .25,fill = "blue", )+annotate("text", x = (27666+27812)/2, y = 8800, label = "ATP8", size = 3) +
  annotate("rect", xmin = 28487, xmax = 29266, ymin = 7500, ymax = 8000,#28487	29266	ATP6
           alpha = .25,fill = "blue")+annotate("text", x = (28487+29266)/2, y = 8200, label = "ATP6", size = 3)+
  annotate("rect", xmin = 36540, xmax = 43647, ymin = 7500, ymax = 8000, #36540	43647	COB
           alpha = .25,fill = "blue")+annotate("text", x = (36540+43647)/2, y = 8300, label = "COB", size = 3)+
  annotate("rect", xmin = 46723, xmax = 46953, ymin = 7500, ymax = 8000, #46723	46953	OLI1
           alpha = .25,fill = "blue", )+annotate("text", x = (46723+46953)/2, y = 8800, label = "OLI1", size = 3) +
  annotate("rect", xmin = 48901, xmax = 50097, ymin = 7500, ymax = 8000, #	VAR1
           alpha = .25,fill = "blue")+annotate("text", x = (48901+50097)/2, y = 8200, label = "VAR1", size = 3)+
  annotate("rect", xmin = 58009, xmax = 62447, ymin = 7500, ymax = 8000, #58009	62447	21S_RRNA 
           alpha = .25,fill = "blue", )+annotate("text", x = (58009+62447)/2, y = 8300, label = "21S rRNA", size = 3) +
  annotate("rect", xmin = 73758, xmax = 74513, ymin = 7500, ymax = 8000, #73758	74513	COX2 
           alpha = .25,fill = "blue", )+annotate("text", x = (74513+73758)/2, y = 8300, label = "COX2", size = 3) +
  annotate("rect", xmin = 79213, xmax = 80022, ymin = 7500, ymax = 8000,#79213	80022	COX3 
           alpha = .25,fill = "blue")+annotate("text", x = (79213+80022)/2, y = 8700, label = "COX3", size = 3)+
  annotate("rect", xmin = 85295, xmax = 85777, ymin = 7500, ymax = 8000,#85295	85777	RPM1
           alpha = .25,fill = "blue", )+annotate("text", x = (85295+85777)/2, y = 8300, label = "RPM1", size = 3)+
  labs(x="", y="")  

```



Plot all mitochondrial coverage lines together
```{r}

mitoc=plot_grid(
  mito_Cu, mito_Mn, mito_Ni, labels = c("","",""),
  ncol = 1, align = "v"
)
#ggsave("figures/mito_coverage.jpeg", width=10, height=6, units = "in", dpi=700)
```



Plot everything together

```{r}

plot_grid(
  both_pet,mitoc,  labels = c("","C"),
  ncol = 1, align = "v", rel_heights = c(1,2.5)
)

plot_grid(
  both_pet,mito_All,  labels = c("","C"),
  ncol = 1, align = "v", rel_heights = c(1,1.5)
)
ggsave("figures/petites_mito_cov.jpeg", width=7, height=8, units = "in", dpi=700)

```





## gene coordinates for mitochondrion
15S_rRNA	chrmt	6546	8194	15S_RRNA 
Q0050	chrmt	13818	16322	AI1 
Q0055	chrmt	13818	18830	AI2 
Q0060	chrmt	13818	19996	AI3 
Q0065	chrmt	13818	21935	AI4 
Q0070	chrmt	13818	23167	AI5_ALPHA 
Q0045	chrmt	13818	26701	COX1 
Q0075	chrmt	24156	25255	AI5_BETA 
Q0080	chrmt	27666	27812	ATP8 
Q0085	chrmt	28487	29266	ATP6 
Q0110	chrmt	36540	38579	BI2 
Q0115	chrmt	36540	40265	BI3 
Q0120	chrmt	36540	42251	BI4 
Q0105	chrmt	36540	43647	COB 
Q0130	chrmt	46723	46953	OLI1 
Q0140	chrmt	48901	50097	VAR1 
21S_rRNA	chrmt	58009	62447	21S_RRNA 
Q0160	chrmt	61022	61729	SCEI 
Q0250	chrmt	73758	74513	COX2 
Q0275	chrmt	79213	80022	COX3 
RPM1	chrmt	85295	85777	RPM1 


