---
title: "8_GO_analysis_PDBupdate"
output: html_document
date: "2024-08-30"
---
```{r}
library(tidyverse)
library(forcats)
library(readr)
library(tidyr)
library(dplyr)
x = getwd()
setwd(x)
options(scipen = 999)
```
Results from SGD GO Term Finder
```{r}
SGD_process <- read_tsv("SGD_process_18Nov2024.tsv")
SGD_process$ANNOTATED_GENES <- gsub("YKR103W", "NFT1", SGD_process$ANNOTATED_GENES)
SGD_process$ANNOTATED_GENES <- gsub("YFL021W", "GAT1", SGD_process$ANNOTATED_GENES)
SGD_function <- read_tsv("SGD_function_18Nov2024.tsv")
SGD_function$ANNOTATED_GENES <- gsub("YKR103W", "NFT1", SGD_function$ANNOTATED_GENES)
SGD_function$ANNOTATED_GENES <- gsub("YFL021W", "GAT1", SGD_function$ANNOTATED_GENES)
SGD_component <- read_tsv("SGD_component_18Nov2024.tsv")
SGD_component$ANNOTATED_GENES <- gsub("YKR103W", "NFT1", SGD_component$ANNOTATED_GENES)
SGD_component$ANNOTATED_GENES <- gsub("YFL021W", "GAT1", SGD_component$ANNOTATED_GENES)
```
Getting lists of genes mapped to each term
```{r}
#Processses
SGD_process$TERM
cellular_process <- SGD_process %>% filter(TERM=="cellular process") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
cellular_component_organization <- SGD_process %>% filter(TERM=="cellular component organization") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
cellular_response_to_stimulus <- SGD_process %>% filter(TERM=="cellular response to stimulus") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
response_to_stimulus <- SGD_process %>% filter(TERM=="response to stimulus") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
biological_regulation <- SGD_process %>% filter(TERM=="biological regulation") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
cellular_component_organization_or_biogenesis <- SGD_process %>% filter(TERM=="cellular component organization or biogenesis") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
regulation_of_cellular_process <- SGD_process %>% filter(TERM=="regulation of cellular process") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
regulation_of_biological_process <- SGD_process %>% filter(TERM=="regulation of biological process") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
cytokinetic_process <- SGD_process %>% filter(TERM=="cytokinetic process") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
polyphosphate_biosynthetic_process <- SGD_process %>% filter(TERM=="polyphosphate biosynthetic process") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))

#Functions
SGD_function$TERM
phosphotransferase_activity <- SGD_function %>% filter(TERM=="phosphotransferase activity, alcohol group as acceptor") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
kinase_activity <- SGD_function %>% filter(TERM=="kinase activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
transferase_activity<- SGD_function %>% filter(TERM=="transferase activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
catalytic_activity <- SGD_function %>% filter(TERM=="catalytic activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
protein_kinase_activity <- SGD_function %>% filter(TERM=="protein kinase activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
ubiquitin_protein_transferase_activity <- SGD_function %>% filter(TERM=="ubiquitin-protein transferase activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
ATP_dependent_activity <- SGD_function %>% filter(TERM=="ATP-dependent activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
catalytic_activity_acting_on_a_protein <- SGD_function %>% filter(TERM=="catalytic activity, acting on a protein") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
MAP_kinase_scaffold_activity <- SGD_function %>% filter(TERM=="MAP-kinase scaffold activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
ubiquitin_ubiquitin_ligase_activity <- SGD_function %>% filter(TERM=="ubiquitin-ubiquitin ligase activity") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))

#Components
SGD_component$TERM
cellular_anatomical_structure <- SGD_component %>% filter(TERM=="cellular anatomical structure") %>% select(ANNOTATED_GENES) %>%
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
intracellular_anatomical_structure <- SGD_component %>% filter(TERM=="intracellular anatomical structure") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
intracellular_organelle <- SGD_component %>% filter(TERM=="intracellular organelle") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
organelle <- SGD_component %>% filter(TERM=="organelle") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
fungal_type_vacuole_membrane <- SGD_component %>% filter(TERM=="fungal-type vacuole membrane") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
lytic_vacuole_membrane <- SGD_component %>% filter(TERM=="lytic vacuole membrane") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
intracellular_membrane_bounded_organelle <- SGD_component %>% filter(TERM=="intracellular membrane-bounded organelle") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
membrane_bounded_organelle <- SGD_component %>% filter(TERM=="membrane-bounded organelle") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
incipient_cellular_bud_site <- SGD_component %>% filter(TERM=="incipient cellular bud site") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
vacuolar_membrane <- SGD_component %>% filter(TERM=="vacuolar membrane") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
cytoplasm <- SGD_component %>% filter(TERM=="cytoplasm") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
vacuolar_transporter_chaperone_complex <- SGD_component %>% filter(TERM=="vacuolar transporter chaperone complex") %>% select(ANNOTATED_GENES) %>% 
  separate_rows(ANNOTATED_GENES, sep = ",") %>% mutate(ANNOTATED_GENES = gsub(' ', '', ANNOTATED_GENES))
```

Checking for redundant terms
If there is greater than 90% overlap in gene lists, then keep the term with the lower pvalue
```{r}
# PROCESS
### TEST: regulation_of_cellular_process and regulation_of_biological_process (69 genes)
inner_join(regulation_of_cellular_process, regulation_of_biological_process) %>% nrow()
### RESULT: 69/69=100% overlap in genes, keep "regulation of cellular process", remove "regulation of biological process"

### TEST: cellular_response_to_stimulus and response_to_stimulus (56 and 59 genes)
inner_join(cellular_response_to_stimulus, response_to_stimulus) %>% nrow()
### RESULT: 56/59=95% overlap in genes, keep "cellular response to stimulus", remove "response to stimulus"

### TEST: cellular_component_organization and cellular_component_organization_or_biogenesis (83 and 92 genes)
inner_join(cellular_component_organization, cellular_component_organization_or_biogenesis) %>% nrow()
### RESULT: 83/92=90.2% overlap in genes, keep "cellular component organization", remove "cellular component organization or biogenesis"



# FUNCTION
### TEST: phosphotransferase_activity and kinase_activity (18 genes)
inner_join(phosphotransferase_activity, kinase_activity) %>% nrow()
### RESULT: 17/18=94% overlap in genes, keep "phosphotransferase activity, alcohol group as acceptor", remove "kinase activity"



# COMPONENT
### TEST: intracellular_anatomical_structure and cellular_anatomical_structure (185 and 193 genes)
inner_join(intracellular_anatomical_structure, cellular_anatomical_structure) %>% nrow()
### RESULT: 185/193=96% overlap in genes, keep "cellular anatomical structure, remove "intracellular anatomical structure"

### TEST: intracellular_organelle and organelle (163 genes)
inner_join(intracellular_organelle, organelle) %>% nrow()
### RESULT: 163/163=100% overlap in genes, keep "intracellular organelle", remove "organelle"

### TEST: intracellular_membrane_bounded_organelle and membrane_bounded_organelle (150 genes)
inner_join(intracellular_membrane_bounded_organelle, membrane_bounded_organelle) %>% nrow()
### RESULT: 150/150=100% overlap in genes, keep "intracellular membrane-bounded organelle", remove "membrane-bounded organelle"

### TEST: intracellular_organelle and intracellular_membrane_bounded_organelle (163 and 150 genes)
inner_join(intracellular_organelle, intracellular_membrane_bounded_organelle) %>% nrow()
### RESULT: 150/163=92% overlap in genes, keep "intracellular membrane-bounded organelle", remove "intracellular organelle"

### TEST: fungal_type_vacuole_membrane and lytic_vacuole_membrane (18 genes)
inner_join(fungal_type_vacuole_membrane, lytic_vacuole_membrane) %>% nrow()
### RESULT: 18/18=100% gene overlap, keep "fungal-type vacuole membrane", remove "lytic vacuole membrane"

### TEST: fungal_type_vacuole_membrane and vacuolar_membrane (18 genes)
inner_join(fungal_type_vacuole_membrane, vacuolar_membrane) %>% nrow()
### RESULT: 18/18=100% gene overlap, keep "fungal-type vacuole membrane", remove "vacuolar membrane"
```


getting colors for FDR representation in tree
```{r}
SGD_function %>% 
  select(TERM, CORRECTED_PVALUE, NUM_LIST_ANNOTATIONS, EXPECTED_FALSE_POSITIVES) %>% 
  ggplot(aes(fct_reorder(TERM, -NUM_LIST_ANNOTATIONS), CORRECTED_PVALUE, color=CORRECTED_PVALUE)) +
  geom_point(size=7) +
  scale_color_gradient(low="goldenrod3", high="lemonchiffon", limits=c(0,0.05)) +
  labs(color='pval', y="pval", x="term") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave("11Nov2024_tree_colors_function.png", height=7, width=6, units = "in")

SGD_process %>% 
  select(TERM, CORRECTED_PVALUE, NUM_LIST_ANNOTATIONS, EXPECTED_FALSE_POSITIVES) %>% 
  ggplot(aes(fct_reorder(TERM, -NUM_LIST_ANNOTATIONS), CORRECTED_PVALUE, color=CORRECTED_PVALUE)) +
  geom_point(size=7) +
  scale_color_gradient(low="magenta4", high="#ffe8fe", limits=c(0,0.05)) +
  labs(color='pval', y="pval", x="term") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave("11Nov2024_tree_colors_process.png", height=7, width=6, units = "in")

SGD_component %>% 
  select(TERM, CORRECTED_PVALUE, NUM_LIST_ANNOTATIONS, EXPECTED_FALSE_POSITIVES) %>% 
  ggplot(aes(fct_reorder(TERM, -NUM_LIST_ANNOTATIONS), CORRECTED_PVALUE, color=CORRECTED_PVALUE)) +
  geom_point(size=7) +
  scale_color_gradient(low="darkcyan", high="lightcyan", limits=c(0,0.05)) +
  labs(color='pval', y="pval", x="term") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave("11Nov2024_tree_colors_component.png", height=7, width=6, units = "in")
```

-------------------------------------------------------------------------------------------------

read in final mutations from Sally's mathematica output
```{r}
filtered_mutations <- read.csv("final_snps_18Nov2024.csv") #this is filtered_mutations from 7_snp_analysis_4Nov2024.Rmd

filtered_mutations$gene <- gsub("YKR103W", "NFT1", filtered_mutations$gene)
filtered_mutations$gene <- gsub("YFL021W", "GAT1", filtered_mutations$gene)
```

#distribution of effects figure
```{r}
#significant terms joining to mutation information

#PROCESS
cellular_process_effect <- cellular_process %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="cellular process")
cellular_component_organization_effect <- cellular_component_organization %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="cellular component organization")
cellular_response_to_stimulus_effect <- cellular_response_to_stimulus %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="cellular response to stimulus")
biological_regulation_effect <- biological_regulation %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="biological regulation")
regulation_of_cellular_process_effect <- regulation_of_cellular_process %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="regulation of cellular process")
cytokinetic_process_effect <- cytokinetic_process %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="cytokinetic process")
polyphosphate_biosynthetic_process_effect <- polyphosphate_biosynthetic_process %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="polyphosphate biosynthetic process")

#FUNCTION
phosphotransferase_activity_effect <- phosphotransferase_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="phosphotransferase activity, alcohol group as acceptor")
transferase_activity_effect <- transferase_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="transferase activity")
catalytic_activity_effect <- catalytic_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="catalytic activity")
protein_kinase_activity_effect <- protein_kinase_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="protein kinase activity")
ubiquitin_protein_transferase_activity_effect <- ubiquitin_protein_transferase_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="ubiquitin-protein transferase activity")
ATP_dependent_activity_effect <- ATP_dependent_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="ATP-dependent activity")
catalytic_activity_acting_on_a_protein_effect <- catalytic_activity_acting_on_a_protein %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="catalytic activity, acting on a protein")
MAP_kinase_scaffold_activity_effect <- MAP_kinase_scaffold_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="MAP-kinase scaffold activity")
ubiquitin_ubiquitin_ligase_activity_effect <- ubiquitin_ubiquitin_ligase_activity %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="ubiquitin-ubiquitin ligase activity")

#COMPONENT
cellular_anatomical_structure_effect <- cellular_anatomical_structure %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="cellular anatomical structure")
fungal_type_vacuole_membrane_effect <- fungal_type_vacuole_membrane %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="fungal-type vacuole membrane")
intracellular_membrane_bounded_organelle_effect <- intracellular_membrane_bounded_organelle %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="intracellular membrane-bounded organelle")
incipient_cellular_bud_site_effect <- incipient_cellular_bud_site %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="incipient cellular bud site")
cytoplasm_effect <- cytoplasm %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="cytoplasm")
vacuolar_transporter_chaperone_complex_effect <- vacuolar_transporter_chaperone_complex %>%
  rename("gene"= ANNOTATED_GENES) %>% 
  left_join(filtered_mutations) %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="vacuolar transporter chaperone complex")
overall <- filtered_mutations %>% 
  group_by(effect) %>% 
  summarize(effect_count=n()) %>% 
  mutate(term="overall")
```
```{r}
#combining terms into one dataframe
effect_term_compare <- rbind(cellular_process_effect, 
                             cellular_component_organization_effect,
                             cellular_response_to_stimulus_effect,
                             biological_regulation_effect,
                             regulation_of_cellular_process_effect, 
                             cytokinetic_process_effect,
                             polyphosphate_biosynthetic_process_effect, 
                             phosphotransferase_activity_effect,
                             transferase_activity_effect,
                             catalytic_activity_effect,
                             protein_kinase_activity_effect,
                             ubiquitin_protein_transferase_activity_effect,
                             ATP_dependent_activity_effect,
                             catalytic_activity_acting_on_a_protein_effect,
                             MAP_kinase_scaffold_activity_effect,
                             ubiquitin_ubiquitin_ligase_activity_effect,
                             cellular_anatomical_structure_effect,
                             fungal_type_vacuole_membrane_effect,
                             intracellular_membrane_bounded_organelle_effect,
                             incipient_cellular_bud_site_effect,
                             cytoplasm_effect,
                             vacuolar_transporter_chaperone_complex_effect,
                             overall)
```
```{r}
effect_term_prop <- effect_term_compare %>% 
  pivot_wider(id_cols=term, names_from = effect, values_from = effect_count) %>%
  mutate_all(funs(replace_na(.,0))) %>% 
  mutate(total_snps = HIGH+MODERATE+LOW) %>% 
  pivot_longer(cols=c(HIGH,MODERATE,LOW), names_to = "effect", values_to = "effect_count") %>% 
  mutate(prop=effect_count/total_snps)

effect_term_prop$effect <- factor(effect_term_prop$effect, levels=c("HIGH", "MODERATE", "LOW"))
effect_term_prop$term <- factor(effect_term_prop$term, 
                                   levels=c("incipient cellular bud site",
                                            "vacuolar transporter chaperone complex",
                                            "fungal-type vacuole membrane",
                                            "intracellular membrane-bounded organelle",
                                            "cytoplasm",
                                            "cellular anatomical structure",
                                            "MAP-kinase scaffold activity",
                                            "ubiquitin-ubiquitin ligase activity",
                                            "ubiquitin-protein transferase activity",
                                            "protein kinase activity",
                                            "phosphotransferase activity, alcohol group as acceptor",
                                            "ATP-dependent activity",
                                            "catalytic activity, acting on a protein",
                                            "transferase activity",
                                            "catalytic activity",
                                            "cytokinetic process",
                                            "polyphosphate biosynthetic process",
                                            "cellular response to stimulus",
                                            "regulation of cellular process",
                                            "biological regulation",
                                            "cellular component organization",
                                            "cellular process",
                                            "overall")) #bottom of graph to top by number of genes
  
effect_term_prop %>% 
  ggplot(aes(x=term, y=prop, fill=effect, width = ifelse(term == "overall", 1, 0.5))) +
  geom_bar(stat="identity") +
  annotate("text", x=1, y=1.05, label="9", size=2.8) + #label is value from total_genes column
  annotate("text", x=2, y=1.05, label="10", size=2.8) +
  annotate("text", x=3, y=1.05, label="37", size=2.8) +
  annotate("text", x=4, y=1.05, label="231", size=2.8) +
  annotate("text", x=5, y=1.05, label="232", size=2.8) +
  annotate("text", x=6, y=1.05, label="294", size=2.8) +
  annotate("text", x=7, y=1.05, label="4", size=2.8) +
  annotate("text", x=8, y=1.05, label="7", size=2.8) +
  annotate("text", x=9, y=1.05, label="17", size=2.8) +
  annotate("text", x=10, y=1.05, label="20", size=2.8) +
  annotate("text", x=11, y=1.05, label="25", size=2.8) +
  annotate("text", x=12, y=1.05, label="32", size=2.8) +
  annotate("text", x=13, y=1.05, label="56", size=2.8) +
  annotate("text", x=14, y=1.05, label="73", size=2.8) +
  annotate("text", x=15, y=1.05, label="123", size=2.8) +
  annotate("text", x=16, y=1.05, label="9", size=2.8) +
  annotate("text", x=17, y=1.05, label="10", size=2.8) +
  annotate("text", x=18, y=1.05, label="87", size=2.8) +
  annotate("text", x=19, y=1.05, label="100", size=2.8) +
  annotate("text", x=20, y=1.05, label="112", size=2.8) +
  annotate("text", x=21, y=1.05, label="135", size=2.8) +
  annotate("text", x=22, y=1.05, label="285", size=2.8) +
  annotate("text", x=23, y=1.05, label="450", size=2.8) +
  scale_fill_manual(limits=c("LOW", "MODERATE", "HIGH"), labels = c("silent", "missense", "nonsense/frameshift"), values=c("#5cd65c", "#ffc60c", "#ff6666"))+
  coord_flip() +
  labs(x="GO Term", y="Proportion", fill="Mutation Effect") +
  theme_bw()
ggsave("GO_distr_of_effects_18Nov2024.png", width=8, height=5, units = "in", dpi=700)
```

#Goodness of Fit test for the above info
```{r}
# All values are taken from effect_term_prop. The order for observed values (from effect_count column) is HIGH, MODERATE, LOW. The expected values are taken from the prop column for term "overall" (order is the same)

expected <- c(0.22666667, 0.61111111, 0.16222222)

observed_cellproc <- c(77,171,37)
observed_cellcomporg <- c(29,84,22)
observed_cellrespstim <- c(17,58,12)
observed_bioreg <- c(27,70,15)
observed_regcellproc <- c(21,64,15)
observed_cytoproc <- c(3,3,3)
observed_polybioproc <- c(9,1,0)
observed_phosphoact <- c(8,12,5)
observed_transact <- c(18,47,8)
observed_catact <- c(25,84,14)
observed_prokinact <- c(6,10,4)
observed_ubiqprotransact <- c(2,14,1)
observed_atpdepact <- c(2,28,2)
observed_catactpro <- c(11,38,7)
observed_mapkinscafact <- c(0,4,0)
observed_ubiqubiqligact <- c(0,7,0)
observed_cellanatstruc <- c(79,179,36)
observed_fungvacmem <- c(19,16,2)
observed_intramemorg <- c(68,135,28)
observed_incipient <- c(1,5,3)
observed_cytoplasm <- c(74,136,22)
observed_vactranschapcomp <- c(9,1,0)


chisq.test(x=observed_cellproc, p=expected)
chisq.test(x=observed_cellcomporg, p=expected)
chisq.test(x=observed_cellrespstim, p=expected)
chisq.test(x=observed_bioreg, p=expected)
chisq.test(x=observed_regcellproc, p=expected)
chisq.test(x=observed_cytoproc, p=expected)
chisq.test(x=observed_polybioproc, p=expected)
chisq.test(x=observed_phosphoact, p=expected)
chisq.test(x=observed_transact, p=expected)
chisq.test(x=observed_catact, p=expected)
chisq.test(x=observed_prokinact, p=expected)
chisq.test(x=observed_ubiqprotransact, p=expected)
chisq.test(x=observed_atpdepact, p=expected)
chisq.test(x=observed_catactpro, p=expected)
chisq.test(x=observed_mapkinscafact, p=expected)
chisq.test(x=observed_ubiqubiqligact, p=expected)
chisq.test(x=observed_cellanatstruc, p=expected)
chisq.test(x=observed_fungvacmem, p=expected)
chisq.test(x=observed_intramemorg, p=expected)
chisq.test(x=observed_incipient, p=expected)
chisq.test(x=observed_cytoplasm, p=expected)
chisq.test(x=observed_vactranschapcomp, p=expected)
```