---
title: "8_GO_analysis_PDBupdate"
output: html_document
date: "2024-08-30"
---
```{r}
library(tidyverse)
library(forcats)
library(readr)
library(tidyr)
library(dplyr)
x = getwd()
setwd(x)
```
Results from PANTHERdb "Functional Classification"
genes taken from snps_FINAL (after removing data from MnBM14 and MnBM42) and uploaded to geneontology.org "GO Enrichment Analysis"
unmapped terms: BUD19, TRT2, YDR269C, YFL021C-A, YLR198C
```{r}
#GO term annotations raw output
all_genes_pdb <- read_tsv("GOterm_analysis/pantherGeneList_final.txt")
```

Wrangling data from pdb output
```{r}
# Split the FUNCTION column into multiple rows based on the semicolon delimiter
# Separate the function descriptions and codes into two columns
function_all <- all_genes_pdb %>%
  dplyr::select(GENE, FUNCTION) %>% 
  separate_rows(FUNCTION, sep = ";") %>%
  mutate(FUNCTION_DESC = sub("\\(GO:\\d+\\)", "", FUNCTION),
         CODE_mf = gsub(".*\\((GO:\\d+)\\).*", "\\1", FUNCTION)) %>%
  dplyr::select(GENE, FUNCTION_DESC)

# Split the PROCESS column into multiple rows based on the semicolon delimiter
# Separate the process descriptions and codes into two columns
process_all <- all_genes_pdb %>%
  dplyr::select(GENE, PROCESS) %>% 
  separate_rows(PROCESS, sep = ";") %>%
  mutate(PROCESS_DESC = sub("\\(GO:\\d+\\)", "", PROCESS),
         CODE_mf = gsub(".*\\((GO:\\d+)\\).*", "\\1", PROCESS)) %>%
  dplyr::select(GENE, PROCESS_DESC)

# Split the COMPONENT column into multiple rows based on the semicolon delimiter
# Separate the component descriptions and codes into two columns
component_all <- all_genes_pdb %>%
  dplyr::select(GENE, COMPONENT) %>% 
  separate_rows(COMPONENT, sep = ";") %>%
  mutate(COMPONENT_DESC = sub("\\(GO:\\d+\\)", "", COMPONENT),
         CODE_mf = gsub(".*\\((GO:\\d+)\\).*", "\\1", COMPONENT)) %>%
  dplyr::select(GENE, COMPONENT_DESC)
```

The Over-representation test groups GO terms hierarchically by parent terms. The ouput from the GO term annotation above has mostly child terms, so I am adding parent terms (gene lists from clicking on parent-terms in over-representation test output page)
```{r}
#load in the parent term gene lists
MF_protein_serine_kinase_activity <- read_tsv("pantherTerm_geneLists/MF_protein_serine_kinase_activity.txt") %>%
  mutate(FUNCTION_DESC="protein serine kinase activity") %>% 
  select(GENE, FUNCTION_DESC)
MF_phosphotransferase_activity <- read_tsv("pantherTerm_geneLists/MF_phosphotransferase_activity.txt") %>% 
  mutate(FUNCTION_DESC="phosphotransferase activity, alcohol group as acceptor") %>% 
  select(GENE, FUNCTION_DESC)
MF_nucleotide_binding <- read_tsv("pantherTerm_geneLists/MF_nucleotide_binding.txt") %>% 
  mutate(FUNCTION_DESC="nucleotide binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_nucleoside_phosphate_binding <- read_tsv("pantherTerm_geneLists/MF_nucleoside_phosphate_binding.txt") %>% 
  mutate(FUNCTION_DESC="nucleoside phosphate binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_kinase_activity <- read_tsv("pantherTerm_geneLists/MF_kinase_activity.txt") %>% 
  mutate(FUNCTION_DESC="kinase activity") %>% 
  select(GENE, FUNCTION_DESC)
MF_ion_binding <- read_tsv("pantherTerm_geneLists/MF_ion_binding.txt") %>% 
  mutate(FUNCTION_DESC="ion binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_heterocyclic_compound_binding <- read_tsv("pantherTerm_geneLists/MF_heterocyclic_compound_binding.txt") %>% 
  mutate(FUNCTION_DESC="heterocyclic compound binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_ATP_binding <- read_tsv("pantherTerm_geneLists/MF_ATP_binding.txt") %>% 
  mutate(FUNCTION_DESC="ATP binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_binding <- read_tsv("pantherTerm_geneLists/MF_binding.txt") %>% 
  mutate(FUNCTION_DESC="binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_anion_binding <- read_tsv("pantherTerm_geneLists/MF_anion_binding_fix.txt") %>% 
  mutate(FUNCTION_DESC="anion binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_small_molecule_binding <- read_tsv("pantherTerm_geneLists/MF_small_molecule_binding.txt") %>% 
  mutate(FUNCTION_DESC="small molecule binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_adenyl_nucleotide_binding <- read_tsv("pantherTerm_geneLists/MF_adenyl_nucleotide_binding.txt") %>% 
  mutate(FUNCTION_DESC="adenyl nucleotide binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_adenyl_ribonucleotide_binding <- read_tsv("pantherTerm_geneLists/MF_adenyl_ribonucleotide_binding.txt") %>% 
  mutate(FUNCTION_DESC="adenyl ribonucleotide binding") %>% 
  select(GENE, FUNCTION_DESC)
MF_catalytic_activity <- read_tsv("pantherTerm_geneLists/MF_catalytic_activity.txt") %>% 
  mutate(FUNCTION_DESC="catalytic activity") %>% 
  select(GENE, FUNCTION_DESC)
CC_site_of_polarized_growth <- read_tsv("pantherTerm_geneLists/CC_site_of_polarized_growth.txt") %>% 
  mutate(COMPONENT_DESC="site of polarized growth") %>% 
  select(GENE, COMPONENT_DESC)
CC_plasma_membrane <- read_tsv("pantherTerm_geneLists/CC_plasma_membrane.txt") %>% 
  mutate(COMPONENT_DESC="plasma membrane") %>% 
  select(GENE, COMPONENT_DESC)
CC_fungal_type_vacuole_membrane <- read_tsv("pantherTerm_geneLists/CC_fungal-type_vacuole_membrane.txt") %>% 
  mutate(COMPONENT_DESC="fungal-type vacuole membrane") %>% 
  select(GENE, COMPONENT_DESC)
CC_cellular_bud_neck <- read_tsv("pantherTerm_geneLists/CC_cellular_bud_neck.txt") %>% 
  mutate(COMPONENT_DESC="cellular bud neck") %>% 
  select(GENE, COMPONENT_DESC)
CC_cell_periphery <- read_tsv("pantherTerm_geneLists/CC_cell_periphery.txt") %>% 
  mutate(COMPONENT_DESC="cell periphery") %>% 
  select(GENE, COMPONENT_DESC)
CC_actin_filament <- read_tsv("pantherTerm_geneLists/CC_actin_filament.txt") %>% 
  mutate(COMPONENT_DESC="actin filament") %>% 
  select(GENE, COMPONENT_DESC)
BP_sporulation <- read_tsv("pantherTerm_geneLists/BP_sporulation.txt") %>% 
  mutate(PROCESS_DESC="sporulation") %>% 
  select(GENE, PROCESS_DESC)
BP_regulation_of_biological_process <- read_tsv("pantherTerm_geneLists/BP_regulation_of_biological_process.txt") %>% 
  mutate(PROCESS_DESC="regulation of biological process") %>% 
  select(GENE, PROCESS_DESC)
BP_protein_polyubiquitination <- read_tsv("pantherTerm_geneLists/BP_protein_polyubiquitination_fix.txt") %>% 
  mutate(PROCESS_DESC="protein polyubiquitination") %>% 
  select(GENE, PROCESS_DESC)
BP_phosphorylation <- read_tsv("pantherTerm_geneLists/BP_phosphorylation.txt") %>% 
  mutate(PROCESS_DESC="phosphorylation") %>% 
  select(GENE, PROCESS_DESC)
BP_cytokinesis <- read_tsv("pantherTerm_geneLists/BP_cytokinesis.txt") %>% 
  mutate(PROCESS_DESC="cytokinesis") %>% 
  select(GENE, PROCESS_DESC)
BP_cellular_response_to_stimulus <- read_tsv("pantherTerm_geneLists/BP_cellular_response_to_stimulus.txt") %>% 
  mutate(PROCESS_DESC="cellular response to stimulus") %>% 
  select(GENE, PROCESS_DESC)
BP_cellular_component_organization <- read_tsv("pantherTerm_geneLists/BP_cellular_component_organization.txt") %>% 
  mutate(PROCESS_DESC="cellular component organization") %>% 
  select(GENE, PROCESS_DESC)
BP_biological_regulation <- read_tsv("pantherTerm_geneLists/BP_biological_regulation.txt") %>% 
  mutate(PROCESS_DESC="biological regulation") %>% 
  select(GENE, PROCESS_DESC)
BP_regulation_of_cellular_process <- read_tsv("pantherTerm_geneLists/BP_regulation_of_cellular_process.txt") %>%
  mutate(PROCESS_DESC="regulation of cellular process") %>% 
  select(GENE, PROCESS_DESC)
BP_cellular_component_organization_or_biogenesis <- read_tsv("pantherTerm_geneLists/BP_cellular_component_organization_or_biogenesis.txt") %>% 
  mutate(PROCESS_DESC="cellular component organization or biogenesis") %>% 
  select(GENE, PROCESS_DESC)
BP_developmental_process <- read_tsv("pantherTerm_geneLists/BP_developmental_process.txt") %>% 
  mutate(PROCESS_DESC="developmental process") %>% 
  select(GENE, PROCESS_DESC)
BP_response_to_stimulus <- read_tsv("pantherTerm_geneLists/BP_response_to_stimulus.txt") %>% 
  mutate(PROCESS_DESC="response to stimulus") %>% 
  select(GENE, PROCESS_DESC)
BP_anatomical_structure_morphogenesis <- read_tsv("pantherTerm_geneLists/BP_anatomical_structure_morphogenesis.txt") %>% 
  mutate(PROCESS_DESC="anatomical structure morphogenesis") %>% 
  select(GENE, PROCESS_DESC)
BP_anatomical_structure_formation_involved_in_morphogenesis <- read_tsv("pantherTerm_geneLists/BP_anatomical_structure_formation_involved_in_morphogenesis.txt") %>% 
  mutate(PROCESS_DESC="anatomical structure formation involved in morphogenesis") %>% 
  select(GENE, PROCESS_DESC)
BP_cellular_process <- read_tsv("pantherTerm_geneLists/BP_cellular_process.txt") %>% 
  mutate(PROCESS_DESC="cellular process") %>% 
  select(GENE, PROCESS_DESC)
BP_cytokinetic_process <- read_tsv("pantherTerm_geneLists/BP_cytokinetic_process_fix.txt") %>% 
  mutate(PROCESS_DESC="cytokinetic process") %>% 
  select(GENE, PROCESS_DESC)
```

```{r}
# join parent term lists to overall term lists and remove duplicates

function_all_fixed <- function_all %>% 
  rbind(MF_protein_serine_kinase_activity,
        MF_phosphotransferase_activity,
        MF_nucleotide_binding,
        MF_nucleoside_phosphate_binding,
        MF_kinase_activity,
        MF_ion_binding,
        MF_heterocyclic_compound_binding,
        MF_ATP_binding,
        MF_catalytic_activity,
        MF_binding,
        MF_small_molecule_binding,
        MF_anion_binding,
        MF_adenyl_nucleotide_binding,
        MF_adenyl_ribonucleotide_binding) %>% 
  distinct()
function_all_fixed %>% write.csv("GOterm_analysis/functionPDB_long.csv")

process_all_fixed <- process_all %>% 
  rbind(BP_sporulation, 
        BP_regulation_of_biological_process, 
        BP_protein_polyubiquitination, 
        BP_phosphorylation, 
        BP_cytokinesis, 
        BP_cellular_response_to_stimulus, 
        BP_cellular_component_organization,
        BP_biological_regulation,
        BP_regulation_of_cellular_process,
        BP_cellular_component_organization_or_biogenesis,
        BP_developmental_process,
        BP_response_to_stimulus,
        BP_anatomical_structure_morphogenesis,
        BP_anatomical_structure_formation_involved_in_morphogenesis,
        BP_cellular_process,
        BP_cytokinetic_process) %>% 
  distinct()
process_all_fixed %>% write.csv("GOterm_analysis/processPDB_long.csv")

component_all_fixed <- component_all %>% 
  rbind(CC_site_of_polarized_growth,
        CC_plasma_membrane,
        CC_fungal_type_vacuole_membrane,
        CC_cellular_bud_neck,
        CC_cell_periphery,
        CC_actin_filament) %>% 
  distinct()
component_all_fixed %>% write.csv("GOterm_analysis/componentPDB_long.csv")
```

Loading in overrepresentation test results to get color gradient for FDR 0-0.05
The files being read in here are overrepresentation output from pantherdb.org
```{r}
overrep_function_pdb <- read_tsv("GOterm_analysis/all_function_Overrepresentation.txt") %>% 
  mutate(FUNCTION_DESC = sub(" \\(GO:\\d+\\)", "", FUNCTION),
         CODE_mf = gsub(".*\\((GO:\\d+)\\).*", "\\1", FUNCTION)) %>%
  select(-c(FUNCTION, `over under`, foldEnrichment))
overrep_function_pdb %>% filter(pval<0.001) %>% write.csv("GOterm_analysis/overrep_p001_function.csv")

overrep_process_pdb <- read_tsv("GOterm_analysis/all_process_Overrepresenation.txt") %>% 
  mutate(PROCESS_DESC = sub(" \\(GO:\\d+\\)", "", PROCESS),
         CODE_mf = gsub(".*\\((GO:\\d+)\\).*", "\\1", PROCESS)) %>% 
  select(-c(PROCESS, `upload_1 (over/under)`, `upload_1 (fold Enrichment)`))
overrep_process_pdb %>% filter(pval<0.001) %>% write.csv("GOterm_analysis/overrep_p001_process.csv")

overrep_component_pdb <- read_tsv("GOterm_analysis/all_component_Overrepresenation.txt") %>% 
  mutate(COMPONENT_DESC = sub(" \\(GO:\\d+\\)", "", COMPONENT),
         CODE_mf = gsub(".*\\((GO:\\d+)\\).*", "\\1", COMPONENT)) %>% 
  select(-c(COMPONENT, `upload_1 (over/under)`, `upload_1 (fold Enrichment)`))
overrep_component_pdb %>% filter(pval<0.001) %>% write.csv("GOterm_analysis/overrep_p001_component.csv")
```

getting colors for FDR representation in tree
```{r}
overrep_function_pdb %>% 
  select(FUNCTION_DESC, List_num, expected, pval, FDR) %>% 
  distinct() %>% 
  filter(FDR<0.05) %>% 
  ggplot(aes(FUNCTION_DESC, pval, color=FDR, size=List_num)) +
  geom_abline(slope=1, intercept = 0) +
  geom_jitter(position = position_dodge(width=0.5)) +
  scale_color_gradient(low="goldenrod3", high="lemonchiffon") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave("GOterm_analysis/figures/tree_colors_function.png")

overrep_process_pdb %>% 
  select(PROCESS_DESC, List_num, expected, pval, FDR) %>% 
  distinct() %>% 
  filter(FDR<0.05) %>% 
  ggplot(aes(PROCESS_DESC, pval, color=FDR, size=List_num)) +
  geom_abline(slope=1, intercept = 0) +
  geom_jitter(position = position_dodge(width=0.5)) +
  scale_color_gradient(low="magenta4", high="#ffe8fe") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave("GOterm_analysis/figures/tree_colors_process.png")

overrep_component_pdb %>% 
  select(COMPONENT_DESC, List_num, expected, pval, FDR) %>% 
  distinct() %>% 
  filter(FDR<0.05) %>% 
  ggplot(aes(COMPONENT_DESC, pval, color=FDR, size=List_num)) +
  geom_abline(slope=1, intercept = 0) +
  geom_jitter(position = position_dodge(width=0.5)) +
  scale_color_gradient(low="darkcyan", high="lightcyan") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
ggsave("GOterm_analysis/figures/tree_colors_component.png")
```
