---
title: "mutation type"
output: html_document
date: '2023-12-28'
---

```{r }
library(tidyverse)
library(ggridges)
library(ggplot2)
library(viridis)
library(cowplot)
#library(hrbrthemes)
x = getwd()
setwd(x)

mut = read.csv("snps/mutation_type.csv")

g = ggplot(mut, aes(mutation))

g + geom_bar(aes(fill = metal)) +
  scale_fill_manual(values=c("#FF9A17", "#3B83FF", "#3EFF3B", "#FD73FF", "white", "black", "#FFF200", "#A880FF"))

table(mut$mutation)

mutDel=mut %>%
  filter(mutation=="del")
mDel = ggplot(mutDel, aes(metal))
mDel + geom_bar(aes(fill = "grey"))
table(mutDel$metal)
```



## mutational spectrum
following categories from the spectrum of single base-pair mutations observed within Gerstein et al. (2015) (8 A:T > G:C, 14 G:C > A:T, 6 A:T > T:A, 12 G:C > T:A, 5 A:T > C:G, 7 G:C > C:G)


```{r }

mut2 = mut %>%
  filter(mutation != "del") %>% #remove deletion
  filter(mutation != "dup")  #remove duplication
mut2$mutation[mut2$mutation == 'A>G'] <- 'A:T > G:C'
mut2$mutation[mut2$mutation == 'T>C'] <- 'A:T > G:C'
mut2$mutation[mut2$mutation == 'G>A'] = "G:C > A:T"
mut2$mutation[mut2$mutation == 'C>T'] = "G:C > A:T"
mut2$mutation[mut2$mutation == "A>T"] = "A:T > T:A"
mut2$mutation[mut2$mutation == "T>A"] = "A:T > T:A"
mut2$mutation[mut2$mutation == "G>T"] = "G:C > T:A"
mut2$mutation[mut2$mutation == "C>A"] = "G:C > T:A"
mut2$mutation[mut2$mutation == "A>C"] = "A:T > C:G"
mut2$mutation[mut2$mutation == "T>G"] = "A:T > C:G"
mut2$mutation[mut2$mutation == "G>C"] = "G:C > C:G"
mut2$mutation[mut2$mutation == "C>G"] = "G:C > C:G"

g = ggplot(mut2, aes(mutation))

g + geom_bar(aes(fill = metal)) +
  scale_fill_manual(values=c("#FF9A17", "#3B83FF", "#3EFF3B", "#FD73FF", "white", "black", "#FFF200", "#A880FF"))


```





set it without MnBM42 and MnBM14

only MnBM42

only MnBM14


```{r }

mutmet = mut2 %>%
  filter(metal!="MnBM42")%>%
  filter(metal!="MnBM14")

altro = ggplot(mutmet, aes(mutation))

altro + geom_bar(aes(fill = metal)) +
  scale_fill_manual(values=c("#FF9A17", "#3B83FF", "#3EFF3B", "#FD73FF", "#FFF200", "#A880FF"))

mutCd=mut2 %>%
  filter(metal=="Cd")
mCd = ggplot(mutCd, aes(mutation))
mCd + geom_bar(aes(fill = "#FF9A17"))

mutCo=mut2 %>%
  filter(metal=="Co")
mCo = ggplot(mutCo, aes(mutation))
mCo + geom_bar(aes(fill = "#3B83FF"))

mutCu=mut2 %>%
  filter(metal=="Cu")
mCu = ggplot(mutCu, aes(mutation))
mCu + geom_bar(aes(fill = "#3EFF3B"))

mutMn=mut2 %>%
  filter(metal=="Mn")
mMn = ggplot(mutMn, aes(mutation))
mMn + geom_bar(aes(fill = "#FD73FF"))

mutNi=mut2 %>%
  filter(metal=="Ni")
mNi = ggplot(mutNi, aes(mutation))
mNi + geom_bar(aes(fill = "#FFF200"))

mutZn=mut2 %>%
  filter(metal=="Zn")
mZn = ggplot(mutZn, aes(mutation))
mZn + geom_bar(aes(fill = "#A880FF"))


mut42 = mut2 %>%
  filter(metal=="MnBM42")
m42 = ggplot(mut42, aes(mutation))
m42 + geom_bar(aes(fill = "hotpink"))

mut14 = mut2 %>%
  filter(metal=="MnBM14")
m14 = ggplot(mut42, aes(mutation, color ="magenta"))
m14 + geom_bar(aes(colour = "magenta"))



```

## Spectrum of base-substitutional mutations from Lynch et al. (2008) Table 1 
Set expectation of proportaions based mutational spectrum from Lynch et al
Summary of studies and their mutation-accumulation study

## Gerstein et al mutational spectrum
Repeat test as sanity check with gerstein data
the spectrum of single base-pair mutations observed within Gerstein et al. (2015) (8 A:T > G:C, 14 G:C > A:T, 6 A:T > T:A, 12 G:C > T:A, 5 A:T > C:G, 7 G:C > C:G)

## spectra of metals
test metals, and the two lines against Lynch et al.
e.g. table(mutCd$mutation)


```{r }


types = c("A:T > G:C", "G:C > A:T", "A:T > T:A", "G:C > T:A", "A:T > C:G", "G:C > C:G")
lynch_other = c(0.102, 0.277, 0.069, 0.313, 0.05, 0.189)
lynch_ma = c(0.118, 0.265, 0.058, 0.265, 0.147, 0.147)
gerstein = c(8, 14, 6, 12, 5, 7)
gerstein.prop = c(8/52, 14/52, 6/52, 12/52, 5/52, 7/52)
Cd = c(8,13,0,2,2,1)
Cd.prop = c(8/26, 13/26, 0/26,2/26,2/26,1/26)
Co = c(9, 25, 12, 13, 9, 12)
Co.prop = c(9/80, 25/80, 12/80, 13/80, 9/80, 12/80)
Cu = c(5, 7, 0, 4, 2, 1)
Cu.prop = c(5/19, 7/19, 0/19, 4/19, 2/19, 1/19)
Mn = c(8, 23, 13, 12, 5, 5)
Mn.prop = c(8/66, 23/66, 13/66, 12/66, 5/66, 5/66)
Ni = c(3, 7, 1, 7, 0, 1)
Ni.prop = c(3/19, 7/19, 1/19, 7/19, 0/19, 1/19)
Zn = c(5, 6, 4, 7, 1, 7)
Zn.prop = c(5/30, 6/30, 4/30, 7/30, 1/30, 7/30)
MnBM42 = c(21, 34, 7, 21, 6, 2)
MnBM42.prop = c(21/91, 34/91, 7/91, 21/91, 6/91, 2/91)
MnBM14 = c(4, 24, 2, 1, 0, 1)
MnBM14.prop = c(4/32, 24/32, 2/32, 1/32, 0/32, 1/32)

spectra = data.frame(types, lynch_other, gerstein, Cd, Co, Cu, Mn, Ni, Zn, MnBM14, MnBM42, gerstein.prop, Cd.prop, Co.prop, Cu.prop, Mn.prop, Ni.prop, Zn.prop, MnBM14.prop, MnBM42.prop)


ger_res <- chisq.test(gerstein, p = lynch_other)
ger_res #X-squared = 7.1557, df = 5, p-value = 0.2093
ger_res1 <- chisq.test(gerstein, p = lynch_ma)
ger_res1 #X-squared = 4.7208, df = 5, p-value = 0.4509


Cd_res <- chisq.test(Cd, p = lynch_other) #X-squared = 25.37, df = 5, p-value = 0.0001182
Co_res <- chisq.test(Co, p = lynch_other) #X-squared = 20.74, df = 5, p-value = 0.0009068
Cu_res <- chisq.test(Cu, p = lynch_other) #X-squared = 10.39, df = 5, p-value = 0.06492
Mn_res <- chisq.test(Mn, p = lynch_other) #X-squared = 26.103, df = 5, p-value = 8.522e-05
Ni_res <- chisq.test(Ni, p = lynch_other) #X-squared = 4.2349, df = 5, p-value = 0.5161
Zn_res <- chisq.test(Zn, p = lynch_other) #X-squared = 4.7585, df = 5, p-value = 0.4461
MnBM42_res <- chisq.test(MnBM42, p = lynch_other) #X-squared = 33.803, df = 5, p-value = 2.606e-06
MnBM14_res <- chisq.test(MnBM14, p = lynch_other) #X-squared = 39.961, df = 5, p-value = 1.521e-07

#bonferroni correction for 8 tests=0.00625


```
Warning: Chi-squared approximation may be incorrect
	Chi-squared test for given probabilities

data:  gerstein
X-squared = 7.1557, df = 5, p-value = 0.2093

Warning: Chi-squared approximation may be incorrect
	Chi-squared test for given probabilities

data:  gerstein
X-squared = 4.7208, df = 5, p-value = 0.4509

> Cd_res

	Chi-squared test for given probabilities

data:  Cd
X-squared = 25.37, df = 5, p-value = 0.0001182

> Co_res

	Chi-squared test for given probabilities

data:  Co
X-squared = 20.74, df = 5, p-value = 0.0009068

> Cu_res

	Chi-squared test for given probabilities

data:  Cu
X-squared = 10.39, df = 5, p-value = 0.06492

> Mn_res

	Chi-squared test for given probabilities

data:  Mn
X-squared = 26.103, df = 5, p-value = 8.522e-05

> Ni_res

	Chi-squared test for given probabilities

data:  Ni
X-squared = 4.2349, df = 5, p-value = 0.5161

> Zn_res

	Chi-squared test for given probabilities

data:  Zn
X-squared = 4.7585, df = 5, p-value = 0.4461

> MnBM42_res

	Chi-squared test for given probabilities

data:  MnBM42
X-squared = 33.803, df = 5, p-value = 2.606e-06

> MnBM14_res

	Chi-squared test for given probabilities

data:  MnBM14
X-squared = 39.961, df = 5, p-value = 1.521e-07




```{r}

pl=ggplot(spectra, aes(x=factor(types), y = lynch_other))+
  geom_bar(stat="identity", width=0.7, fill="grey")+
  theme_minimal() +
  ylim(0,0.75) +
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="Lynch et al. (2008)")

pg=ggplot(spectra, aes(x=factor(types), y = gerstein.prop))+
  geom_bar(stat="identity", width=0.7, fill="black")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="Gerstein et al. (2015)")
pcd=ggplot(spectra, aes(x=factor(types), y = Cd.prop))+
  geom_bar(stat="identity", width=0.7, fill="#FF9A17")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="*Cadmium")
pco=ggplot(spectra, aes(x=factor(types), y = Co.prop))+
  geom_bar(stat="identity", width=0.7, fill="#3B83FF")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="*Cobalt")
pcu=ggplot(spectra, aes(x=factor(types), y = Cu.prop))+
  geom_bar(stat="identity", width=0.7, fill="#3EFF3B")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="Copper")
pmn=ggplot(spectra, aes(x=factor(types), y = Mn.prop))+
  geom_bar(stat="identity", width=0.7, fill="#FD73FF")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="*Manganese (ex. MnBM14 & MnBM42)")
pni=ggplot(spectra, aes(x=factor(types), y = Ni.prop))+
  geom_bar(stat="identity", width=0.7, fill="#FFF200")+
  theme_minimal()+
  ylim(0,0.7)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="Nickel")
pzn=ggplot(spectra, aes(x=factor(types), y = Zn.prop))+
  geom_bar(stat="identity", width=0.7, fill="#A880FF")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="Zinc")
p14=ggplot(spectra, aes(x=factor(types), y = MnBM14.prop))+
  geom_bar(stat="identity", width=0.7, fill="pink")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="*MnBM14")
p42=ggplot(spectra, aes(x=factor(types), y = MnBM42.prop))+
  geom_bar(stat="identity", width=0.7, fill="hotpink")+
  theme_minimal()+
  ylim(0,0.75)+
  theme(axis.text.x  = element_text(size=10))+
  theme(axis.text.y  = element_text(size=10))+
  labs(x=NULL, y=NULL, title="*MnBM42")


```

##plot proportions of the different mutational spectra
```{r}

plot_grid(pl,pg,pcd, pco,pcu,pmn,pni,pzn,p14,p42, labels = c("A","B","C","D","E","F","G","H","I","J"), ncol = 2, align = "h")

#ggsave("mutationtypes_suppl.jpeg", width=12, height=14, units = "in", dpi=700)
ggsave("figures/mutationtypes_suppl.jpeg", width=12, height=14, units = "in", dpi=700)

```


