---
title: "Metal resistance genes"
output: html_document
---
#Setup
##Load libraries
```{r}
library(tidyverse)
library(stringr)
```

##Set working directory
```{r}
x = getwd()
setwd(x)
```

##Read in file
Read in data taken directly from the metal resistance observable phenotype page on SGD
```{r}
genes <- read_tsv("metal_resistance_annotations_2025.txt")
metal_resistance_genes <- genes$Gene %>% unique()
```
#Data cleaning

##Filter for our metals
Extract metal information from "Chemical" column and filter for metals of interest
```{r}
genes3 <- genes %>% 
  mutate(Metal = case_when(grepl("copper", Chemical, ignore.case = TRUE) ~"copper",
                            grepl("cadmi", Chemical, ignore.case = TRUE) ~"cadmium",
                            grepl("nickel", Chemical, ignore.case = TRUE) ~"nickel",
                            grepl("zinc", Chemical, ignore.case = TRUE) ~"zinc",
                            grepl("manga", Chemical, ignore.case = TRUE) ~"manganese",
                            grepl("coba", Chemical, ignore.case = TRUE) ~"cobalt",
                            )) %>% 
  distinct() %>% 
  filter(Metal == "cadmium" | 
         Metal == "cobalt" |
         Metal == "copper" | 
         Metal == "manganese" | 
         Metal == "nickel" |
         Metal == "zinc") %>% 
  select(Gene, Phenotype, 'Mutant Information', Metal, Chemical) %>% 
  rename("Mutant.Information" = 'Mutant Information')
```

##Assign mutation effect
Evaluate effect of mutation (null vs. overexpression) on phenotype (increase vs. decrease resistance)
```{r}
synergism <- genes3 %>% 
  mutate(activity.sign = case_when(grepl("null", Mutant.Information, ignore.case = TRUE) ~"-",
                                   grepl("overexp", Mutant.Information, ignore.case = TRUE) ~"+")) %>%
  mutate(resist.sign = case_when(grepl("increa", Phenotype, ignore.case = TRUE) ~"+",
                                   grepl("decrea", Phenotype, ignore.case = TRUE) ~"-")) %>%
  mutate(product.sign = ifelse(activity.sign == resist.sign, "+",
                        ifelse(activity.sign != resist.sign, "-", NA))) %>% 
  drop_na()
```

##Make list for each metal
Form lists of mutation effects for each metal
```{r}
cadmium_syn <- synergism %>% 
  filter(Metal=="cadmium")
cobalt_syn <- synergism %>% 
  filter(Metal=="cobalt")
copper_syn <- synergism %>% 
  filter(Metal=="copper")
manganese_syn <- synergism %>% 
  filter(Metal=="manganese")
nickel_syn <- synergism %>% 
  filter(Metal=="nickel")
zinc_syn <- synergism %>% 
  filter(Metal=="zinc")
```

##Determine overlap in mutation effects
Merge lists for metal pairs and omit NAs so only genes in common remain. Determine if the two metals agree on the effect of a mutation on resistance. Calculate proportion of agreements in gene set and number of genes in the set.
```{r}
CdCo <- merge(cadmium_syn,cobalt_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>% 
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CdCu <- merge(cadmium_syn,copper_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CdMn <- merge(cadmium_syn,manganese_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CdNi <- merge(cadmium_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CdZn <- merge(cadmium_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CoCu <- merge(cobalt_syn,copper_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CoMn <- merge(cobalt_syn,manganese_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CoNi <- merge(cobalt_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CoZn <- merge(cobalt_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CuMn <- merge(copper_syn,manganese_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

CuNi <- merge(copper_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CuNi$pos.pleio)/nrow(CuNi)
nrow(CuNi)

CuZn <- merge(copper_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

MnNi <- merge(manganese_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

MnZn <- merge(manganese_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))

NiZn <- merge(nickel_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
```

##Calculate values for overlap
```{r}
sum(CdCo$pos.pleio)/nrow(CdCo)
nrow(CdCo)
sum(CdCu$pos.pleio)/nrow(CdCu)
nrow(CdCu)
sum(CdMn$pos.pleio)/nrow(CdMn)
nrow(CdMn)
sum(CdNi$pos.pleio)/nrow(CdNi)
nrow(CdNi)
sum(CdZn$pos.pleio)/nrow(CdZn)
nrow(CdZn)
sum(CoCu$pos.pleio)/nrow(CoCu)
nrow(CoCu)
sum(CoMn$pos.pleio)/nrow(CoMn)
nrow(CoMn)
sum(CoNi$pos.pleio)/nrow(CoNi)
nrow(CoNi)
sum(CoZn$pos.pleio)/nrow(CoZn)
nrow(CoZn)
sum(CuMn$pos.pleio)/nrow(CuMn)
nrow(CuMn)
sum(CuNi$pos.pleio)/nrow(CuNi)
nrow(CuNi)
sum(CuZn$pos.pleio)/nrow(CuZn)
nrow(CuZn)
sum(MnNi$pos.pleio)/nrow(MnNi)
nrow(MnNi)
sum(MnZn$pos.pleio)/nrow(MnZn)
nrow(MnZn)
sum(NiZn$pos.pleio)/nrow(NiZn)
nrow(NiZn)
```

##Put calculated values into dataframe
Make dataframe with values from above
```{r}
prop_pleio <- data.frame(
  combo <- c("CdCo","CdCu","CdMn","CdNi","CdZn","CoCu","CoMn","CoNi","CoZn","CuMn","CuNi","CuZn","MnNi","MnZn","NiZn"),
  met1 <- c("Cd","Cd","Cd","Cd","Cd","Co","Co","Co","Co","Cu","Cu","Cu","Mn","Mn","Ni"),
  met2 <- c("Co","Cu","Mn","Ni","Zn","Cu","Mn","Ni","Zn","Mn","Ni","Zn","Ni","Zn","Zn"),
  prop.syn <- c(0.8503401,0.8564593,0.8691275,0.6518987,0.7252252,0.44,0.8396226,0.8484848,0.9323308,0.9192825,0.5714286,0.4305556,0.7173913,0.6358025,0.8188976),
  no.genes <- c(147,209,298,158,222,25,106,132,133,223,21,72,92,162,127)
) 
prop_pleio %>% write.csv("prop_pleio_2025.csv")
```

#Visualize data
```{r}
prop_pleio %>% 
  ggplot(aes(x=met1,y=met2,color=prop.syn)) + 
  geom_point(aes(size = no.genes), shape=15) +
  scale_color_gradient(low = "yellow", high = "blue",
  name="Proportion\npositive\npleiotropy") +
  scale_x_discrete(position = "top") + 
  scale_size(range=c(8,18), name="Number of genes\nin common", breaks=c(50,100,200)) +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank(),
        axis.text = element_text(size=12))
ggsave("SGD_predictions.jpeg", width=7, height=5, units = "in", dpi=900)
```
