---
title: "Metal resistance genes"
output: html_document
header-includes: \usepackage{colortbl}
---

```{r}
library(tidyverse)
library(stringr)
library(VennDiagram)
library(gtools)
library(data.table)
library(viridis)
library(formattable)
```

Read in data taken directly from the metal resistance observable phenotype page on SGD

```{r}
x = getwd()
setwd(x)

genes <- read.csv("metal_resistance_annotations.csv") #This file is from Aug 24, 2021 https://www.yeastgenome.org/observable/APO:0000090
```

Extract metal information from "Chemical" column and filter for metals of interest

```{r}
genes3 <- genes %>% 
  mutate(Metal = case_when(grepl("copper", Chemical, ignore.case = TRUE) ~"copper",
                            grepl("cadmi", Chemical, ignore.case = TRUE) ~"cadmium",
                            grepl("nickel", Chemical, ignore.case = TRUE) ~"nickel",
                            grepl("zinc", Chemical, ignore.case = TRUE) ~"zinc",
                            grepl("arsen", Chemical, ignore.case = TRUE) ~"arsenic",
                            grepl("gold", Chemical, ignore.case = TRUE) ~"gold",
                            grepl("manga", Chemical, ignore.case = TRUE) ~"manganese",
                            grepl("merc", Chemical, ignore.case = TRUE) ~"mercury",
                            grepl("coba", Chemical, ignore.case = TRUE) ~"cobalt",
                            grepl("alumi", Chemical, ignore.case = TRUE) ~"aluminum",
                            grepl("iron", Chemical, ignore.case = TRUE) ~"iron",
                            grepl("selen", Chemical, ignore.case = TRUE) ~"calcium",
                            grepl("tin", Chemical, ignore.case = TRUE) ~"tin",
                            grepl("chrom", Chemical, ignore.case = TRUE) ~"chromium",
                            grepl("lead", Chemical, ignore.case = TRUE) ~"lead"
                            )) %>% 
  na.omit() %>% 
  distinct() %>% 
  filter(Metal == "cadmium" | 
         Metal == "cobalt" |
         Metal == "copper" | 
         Metal == "manganese" | 
         Metal == "nickel" |
         Metal == "zinc") %>% 
  select(Gene, Phenotype, Mutant.Information, Metal, Chemical)
```

Evaluate effect of mutation (null vs. overexpression) on phenotype (increase vs. decrease resistance)

```{r}
synergism <- genes3 %>% 
  mutate(activity.sign = case_when(grepl("null", Mutant.Information, ignore.case = TRUE) ~"-",
                                   grepl("overexp", Mutant.Information, ignore.case = TRUE) ~"+")) %>%
  mutate(resist.sign = case_when(grepl("increa", Phenotype, ignore.case = TRUE) ~"+",
                                   grepl("decrea", Phenotype, ignore.case = TRUE) ~"-")) %>%
  mutate(product.sign = ifelse(activity.sign == resist.sign, "+",
                        ifelse(activity.sign != resist.sign, "-", NA)))
```

Form lists of mutation effects for each metal

```{r}
cadmium_syn <- synergism %>% 
  filter(Metal=="cadmium")
cobalt_syn <- synergism %>% 
  filter(Metal=="cobalt")
copper_syn <- synergism %>% 
  filter(Metal=="copper")
manganese_syn <- synergism %>% 
  filter(Metal=="manganese")
nickel_syn <- synergism %>% 
  filter(Metal=="nickel")
zinc_syn <- synergism %>% 
  filter(Metal=="zinc")
```

Merge lists for metal pairs and omit NAs so only genes in common remain. Determine if the two metals agree on the effect of a mutation on resistance. Calculate proportion of agreements in gene set and number of genes in the set.

```{r}
CdCo <- merge(cadmium_syn,cobalt_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>% 
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CdCo$pos.pleio)/nrow(CdCo)
nrow(CdCo)
```

```{r}
CdCu <- merge(cadmium_syn,copper_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CdCu$pos.pleio)/nrow(CdCu)
nrow(CdCu)
```

```{r}
CdMn <- merge(cadmium_syn,manganese_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CdMn$pos.pleio)/nrow(CdMn)
nrow(CdMn)
```

```{r}
CdNi <- merge(cadmium_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CdNi$pos.pleio)/nrow(CdNi)
nrow(CdNi)
```

```{r}
CdZn <- merge(cadmium_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CdZn$pos.pleio)/nrow(CdZn)
nrow(CdZn)
```

```{r}
CoCu <- merge(cobalt_syn,copper_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CoCu$pos.pleio)/nrow(CoCu)
nrow(CoCu)
```

```{r}
CoMn <- merge(cobalt_syn,manganese_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CoMn$pos.pleio)/nrow(CoMn)
nrow(CoMn)
```

```{r}
CoNi <- merge(cobalt_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CoNi$pos.pleio)/nrow(CoNi)
nrow(CoNi)
```

```{r}
CoZn <- merge(cobalt_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CoZn$pos.pleio)/nrow(CoZn)
nrow(CoZn)
```

```{r}
CuMn <- merge(copper_syn,manganese_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CuMn$pos.pleio)/nrow(CuMn)
nrow(CuMn)
```

```{r}
CuNi <- merge(copper_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CuNi$pos.pleio)/nrow(CuNi)
nrow(CuNi)
```

```{r}
CuZn <- merge(copper_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(CuZn$pos.pleio)/nrow(CuZn)
nrow(CuZn)
```

```{r}
MnNi <- merge(manganese_syn,nickel_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(MnNi$pos.pleio)/nrow(MnNi)
nrow(MnNi)
```

```{r}
MnZn <- merge(manganese_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(MnZn$pos.pleio)/nrow(MnZn)
nrow(MnZn)
```

```{r}
NiZn <- merge(nickel_syn,zinc_syn,by="Gene") %>% 
  select(Gene, Metal.x, Metal.y, product.sign.x, product.sign.y) %>% 
  distinct() %>%  
  na.omit() %>% 
  mutate(pos.pleio = case_when(product.sign.x==product.sign.y ~ 1,
                               product.sign.x!=product.sign.y ~ 0))
sum(NiZn$pos.pleio)/nrow(NiZn)
nrow(NiZn)
```

Make dataframe with values from above

```{r}
combo <- c("CdCo","CdCu","CdMn","CdNi","CdZn","CoCu","CoMn","CoNi","CoZn","CuMn","CuNi","CuZn","MnNi","MnZn","NiZn")
met1 <- c("Cd","Cd","Cd","Cd","Cd","Co","Co","Co","Co","Cu","Cu","Cu","Mn","Mn","Ni")
met2 <- c("Co","Cu","Mn","Ni","Zn","Cu","Mn","Ni","Zn","Mn","Ni","Zn","Ni","Zn","Zn")
prop.syn <- c(0.8482759,0.8627451,0.8754325,0.6558442,0.7239819,0.44,0.8446602,0.8615385,0.9318182,0.9276018,0.5789474,0.4225352,0.744186,0.6329114,0.8253968)
no.genes <- c(145,204,289,154,221,25,103,130,132,221,19,71,86,158,126)

prop_pleio <- data.frame(combo, met1, met2, prop.syn, no.genes)
```

Visualize data

```{r}
prop_pleio %>% 
  ggplot(aes(x=met1,y=met2,color=prop.syn)) + 
    geom_point(aes(size = no.genes), shape=15) +
    scale_color_gradient(low = "yellow", high = "blue",
    name="Proportion\npositive\npleiotropy") +
    scale_x_discrete(position = "top") + 
    theme_bw() +
    theme(axis.title.x=element_blank(), axis.title.y=element_blank()) +
    scale_size(range=c(8,18), name="Number of genes\nin common")
ggsave("SGD_predictions.jpeg", width=7, height=5, units = "in", dpi=900)
```
